\begin{thebibliography}{00}
\bibitem{ji2023survey} Z. Ji et al., "Survey of Hallucination in Natural Language Generation," \textit{ACM Computing Surveys}, vol. 55, no. 12, pp. 1--38, 2023.
\bibitem{manakul2023selfcheckgpt} P. Manakul et al., "SelfCheckGPT: Zero-Resource Black-Box Hallucination Detection for Generative Large Language Models," \textit{EMNLP}, 2023.
\bibitem{kuhn2023semantic} L. Kuhn et al., "Semantic Uncertainty: Linguistic Invariances for Uncertainty Estimation in Natural Language Generation," \textit{ICLR}, 2023.
\bibitem{elhage2021mathematical} N. Elhage et al., "A Mathematical Framework for Transformer Circuits," \textit{Anthropic}, 2021.
\bibitem{olah2020zoom} C. Olah et al., "Zoom In: An Introduction to Circuits," \textit{Distill}, 2020.
\bibitem{olsson2022context} C. Olsson et al., "In-context Learning and Induction Heads," \textit{Transformer Circuits Thread}, 2022.
\bibitem{meng2022locating} K. Meng et al., "Locating and Editing Factual Associations in GPT," \textit{NeurIPS}, 2022.
\bibitem{tenney2019bert} I. Tenney et al., "BERT Rediscovers the Classical NLP Pipeline," \textit{ACL}, 2019.
\bibitem{geva2021transformer} M. Geva et al., "Transformer Feed-Forward Layers Are Key-Value Memories," \textit{EMNLP}, 2021.
\bibitem{elhage2022toy} N. Elhage et al., "Toy Models of Superposition," \textit{Anthropic}, 2022.
\bibitem{bills2023language} S. Bills et al., "Language Models Can Explain Neurons in Language Models," \textit{OpenAI}, 2023.
\bibitem{nanda2022transformerlens} N. Nanda et al., "TransformerLens," \textit{GitHub}, 2022.
\end{thebibliography}

